### 2013-12-03 docs/paper/fsuthesis/sample/citations.tex
docs/paper/fsuthesis/thesis-template/paper.tex has my paper

%%% For this chapter, I need a few extra definitions.  First, I create a
%%% new macro so that I can easily type the BibTeX logo.

\def\BibTeX{{\rmfamily B\kern-.05em%
   \textsc{i\kern-.025em b}\kern-.08em\TeX}}

%%% Next I create a new environment called 'narrowref' simulating the
%%% appearance of the 'references' environment.  The margins are
%%% narrowed to set off the material a little better.  I also invoke
%%% the 'singlespaced' environment to demonstrate the look of the
%%% actual output of the references in case I'm processing this
%%% document using 'expanded' mode.  If you'd like to create your own
%%% environments for special purposes, refer to the references for the
%%% details.  For aspiring LaTeX experts, note that I reset
%%% the \parskip parameter *after* beginning the first paragraph
%%% with \leavevmode.  This is because I only want to use the new
%%% paragraph spacing between paragraphs contained *within* this
%%% environment, and the default paragraph spacing outside the
%%% environment.  (The \parskip glue is added before the start of a
%%% new paragraph, not at the end of the old paragraph.)

\newenvironment{narrowref}{
  \begin{singlespaced}
  \frenchspacing
  \setlength{\leftskip}{4em}
  \setlength{\rightskip}{3em}
  \setlength{\parindent}{-2em}
  \addvspace{0.5\baselineskip}
  \leavevmode
  \setlength{\parskip}{0.6\baselineskip}\ignorespaces}
 {\end{singlespaced}\addvspace{.5\baselineskip}}

%%% Now let the chapter begin.

\chapter{Citations and References}
In this chapter I demonstrate how to use features of \LaTeX{} and
the \pkg{fsuthesis} class to help me create a list of references or
bibliography.  There are two options: (1)~use the \lit{references}
environment and enter bibliographic entries with styling 
that I create myself, or (2)~use the \LaTeX\ \verb|\cite| macro
and \BibTeX{} to order and style the bibliographic entries according
to a pre-defined style sheet.

\section{Using the \texttt{references} Environment}
Let's start simply.  Assuming we don't have many citations, we'll
create a references section manually using the
\verb|\begin{references}| environment provided by the
\textsf{fsuthesis} class.  If your discipline has a style
guide for presenting references, you should use that information to
create your own entries.  Here are some example entries with
explicit formatting specified:
\begin{verbatim}
\begin{references}
Picaut, J., F. Masia, and Y. du Penhoat, 1997: An advective-reflective
conceptual model for the oscillatory nature of the ENSO.
\textit{Science}, \textbf{277}, 663--666.

Yasunari, T., 1990: Impact of Indian monsoon on the coupled
atmosphere/ocean system in the tropical Pacific.
\textit{Meteor. Atmos. Phys.}, \textbf{44}, 19--41.
\end{references}
\end{verbatim}

Note that I had to specify italics and bold-facing myself, according
to the style guide that I'm using.  (If you don't have a
discipline-specific style guide, see \cite{Anonymous:1993:CMS}
or \cite{Anonymous:2009:PMA} for lots of bibliographic examples.)
The \lit{references} environment provides exdented entries, spacing,
and a heading, but the rest of the formatting is up to you.  Likewise,
citations of these references within your document must be formatted
manually.  Once processed, the reference entries above look like the
following:
\begin{narrowref}
Picaut, J., F. Masia, and Y. du Penhoat, 1997: An advective-reflective
conceptual model for the oscillatory nature of the ENSO.
\textit{Science}, \textbf{277}, 663--666.

Yasunari, T., 1990: Impact of Indian monsoon on the coupled
atmosphere/ocean system in the tropical Pacific.
\textit{Meteor. Atmos. Phys.}, \textbf{44}, 19--41.
\end{narrowref}

%%% My \BibTeX macro fails me in this section title, because there are
%%% no smallcaps bold-faced fonts provided by default with my LaTeX
%%% installation.  For this one place, I do something slightly different.

\section{Citations and B{\small IB}\TeX}
If your thesis or dissertation does not have many citations or
references, then the \lit{references} environment may be all you
need. However, if you have more than a handful of citations, you owe
it to yourself to invest a little more energy into learning about the
powerful, time-saving features of \BibTeX.  To use \BibTeX,
bibliographic entries are added to an external file with attributes
identifying elements of the entry, such as authors, titles, journals,
etc.  You may then cite a reference in your document using its unique
key.  Many disciplines have developed large \BibTeX{} databases
already, so if you're lucky, you only need download a pre-built file
ready to go.  You can always add a few more references if those
entries don't already exist in the file you download.

For this sample document, I have created a small \BibTeX{}
bibliography database in a file called \lit{myrefs.bib} which is
excerpted from a larger collection of pre-generated \TeX-related
entries I downloaded from the web.  If your discipline does not
already distribute \BibTeX{} databases publicly, it's just a bit
more typing to create your own \BibTeX{} file.  As an example, here's
what a \BibTeX{} entry might look like:
\begin{verbatim}
@Article{Picaut:1997,                  
  author = "J. Picaut and F. Masia and Y. du Penhoat",
  title = "An advective-reflective conceptual model for
    the oscillatory nature of the ENSO",
  volume = "277",
  year = "1997",
  journal = "Science",
  pages = "663--666"
}
\end{verbatim}

Another advantage of the \BibTeX{} approach is that you can continue
to add to this database throughout your professional career, creating
entries as you read books and journal articles you may want to
reference in the future.  I encourage you to refer to the
standard \LaTeX{} references (e.g., \cite{Lamport:1994:LDP}
and \cite{Kopka:2004:GLT}) or to search the web for further
information on creating your own \BibTeX{} database.

Once you've created or downloaded a \BibTeX{} database, you may cite a
document using its unique \textit{key} as an argument to
the \LaTeX{} \verb|\cite| macro.  For example, in the previous
paragraph, I used the commands \verb|\cite{Lamport:1994:LDP}| and
\verb|\cite{Kopka:2004:GLT}|, where \lit{Lamport:1994:LDP} and
\lit{Kopka:2004:GLT} are the keys for their respective documents.
(The keys are defined in the \BibTeX\ database file.)  By making these
citations, the bibliographic entries for these documents will be
pulled from my \BibTeX{} database and added to the bibliography
automatically.

I can also add entries to the bibliography without having citations in
my text by using the \verb|\nocite{key}| command with the
desired \textit{key}.  If I want every entry in my \BibTeX{} file
inserted into my bibliography, then I can issue the
command \verb|\nocite{*}| as a shortcut.  (You don't have to worry
about citations being redundant: bibliographic entries will be
included only once no matter how many times they may be \verb|\cite|d
or \verb|\nocite|d in your document.)

\section{Alternative Citation and Bibliography Formats}
The pre-formatted version of this document uses the default citation
and bibliography formatting supplied by \LaTeX\ and \BibTeX.  There
are a few options to alter the appearance of citations and the
bibliography.  (See one of the references for your choices.)  However,
many disciplines and journals have created their own
advanced \BibTeX{} formatting styles, and you may want to use one of
these in your document.

\subsection{The \pkg{natbib} Package}
The \pkg{natbib} package, activated by
including \verb|\usepackage{natbib}| in the document preamble,
provides a common alternative to the default \LaTeX\ bibliography and
citation styles.  Among other things, this package will allow you to
make author--year citations in your document automatically (e.g.,
``Fargunkle et al., (2001)'') without any changes to the \BibTeX{}
database file.  In addition to the default \LaTeX\ \verb|\cite| macro,
the package provides several new macros for creating citations in
different contexts.

This document's driver file (\lit{thesis.tex}) has options for trying
out the \pkg{natbib} package.  Look for the \verb|\usepackage{natbib}|
line near the top of the file, and then scroll down to the
bibliography section for additional style selections you must specify
for the bibliography itself.  For a bibliography formatted according
to the style of the Association for Computing Machinery, for example,
you would issue the command \verb|\bibliographystyle{acm}| before
the \verb|\bibliography| command.  See \cite{Kopka:2004:GLT} (or
search the web) for more information on \pkg{natbib}.

\subsection{The \pkg{apacite} Package}
The APA (American Psychological Association) created another
commonly-used style guide for citations and bibliographic formatting.
The \LaTeX\ package \pkg{apacite} was designed to follow the APA
guidelines.  The package provides its own set of macros for specifying
citations in the document text, but it also has a mode for
compatibility with the popular \pkg{natbib} package.  With this
compatibility mode enabled, you may use \pkg{natbib}-style citation
commands in your document to generate the APA-style citations and
bibliography.  In addition, using this compatibility mode means that
you may switch between these two package alternatives without having
to re-write the citations in your document.

If you use the \pkg{apacite} package, and if you have also enabled
the \pkg{hyperref} package for your document, you \emph{must} enable
\lit{natbibapa} compatibility mode, or citations will give you
cryptic error messages when your document is processed.  Load the
package using the following line:
\begin{verbatim}
\usepackage[natbibapa]{apacite}
\end{verbatim}
The corresponding bibliography style would be set by the following 
command line:
\begin{verbatim}
\bibliographystyle{apacite}
\end{verbatim}

\section{Even More Options}

For those writing documents in the humanities or history disciplines,
there are several other options you may want to consider.  Because
of limitations in the older \BibTeX{} system, there is a move toward a
newer system called \pkg{biblatex}.  Bibliography styles based on this
system use different citation commands and generate bibliographies
in a different way.

For a comparison of the \BibTeX{} and \pkg{biblatex} systems, the
links below provide useful information.  To use the \pkg{biblatex}
package, you will have to customize the \textsf{fsuthesis} class to
accommodate.
\begin{itemize}\small
\item \textsf{http://tex.stackexchange.com/questions/25701/bibtex-vs-biber-and-biblatex-vs-natbib}
\item \textsf{http://tex.stackexchange.com/questions/5091/what-to-do-to-switch-to-biblatex}
\end{itemize}

\subsection{The \pkg{biblatex-chicago} Package}
While the \pkg{natbib} package provides an optional bibliography style
that closely follows the so-called ``Chicago style''
(see \cite{Anonymous:1993:CMS}), the \pkg{biblatex-chicago} package
provides many more features and stricter stylistic adherence over a
wider range of types of references.

This package makes use of a new ``back-end'' processor
called \lit{biber} rather than the \BibTeX{} program.  The
major \TeX{} distributions already include the required programs.
However, you may need to configure your environment to run \lit{biber}
rather than \BibTeX{} to process your document.

\subsection{The \pkg{biblatex-historian} Package}
A \pkg{historian} package is similar to \pkg{biblatex-chicago}, but it
is tuned for documents that require extensive footnote citations in the
Turabian/Chicago styles.  The user manual provides detailed
descriptions of fields, citation commands, and plenty of examples.
This package also requires the \lit{biber} back-end processor.