### 2013-12-03 docs/paper/fsuthesis/sample/history.tex

\chapter{The World According to Student~Bloopers}

One of the fringe benefits of being an English or History teacher is
receiving the occasional jewel of a student blooper in an essay. I
have pasted together the following ``history'' of the world from
certifiably genuine student bloopers collected by teachers throughout
the United States, from eight grade through college level. Read
carefully, and you will learn a lot.

The inhabitants of Egypt were called mummies. They lived in the Sarah
Dessert and traveled by Camelot. The climate of the Sarah is such that
the inhabitants have to live elsewhere, so certain areas of the
dessert are cultivated by irritation. The Egyptians built the
Pyramids in the shape of a huge triangular cube. The Pramids are a
range of mountains between France and Spain.

The Bible is full of interesting caricatures. In the first book of the
Bible, Guinesses, Adam and Eve were created from an apple tree. One of
their children, Cain, asked ``Am I my brother's son?'' God asked Abraham
to sacrifice Issac on Mount Montezuma. Jacob, son of Issac, stole his
brother's birthmark. Jacob was a partiarch who brought up his twelve
sons to be partiarchs, but they did not take to it. One of Jacob's
sons, Joseph, gave refuse to the Israelites.

Pharaoh forced the Hebrew slaves to make bread without straw. Moses
led them to the Red Sea, where they made unleavened bread, which is
bread made without any ingredients. Afterwards, Moses went up on Mount
Cyanide to get the ten commandments. David was a Hebrew king skilled
at playing the liar. He fougth with the Philatelists, a race of people
who lived in Biblical times. Solomon, one of David's sons, had 500
wives and 500 porcupines.

Without the Greeks, we wouldn't have history. The Greeks invented
three kinds of columns---Corinthian, Doric and Ironic. They also had
myths. A myth is a female moth. One myth says that the mother of
Achilles dipped him in the River Stynx until he became
intolerable. Achilles appears in ``The Illiad'', by Homer. Homer also
wrote the ``Oddity'', in which Penelope was the last hardship that
Ulysses endured on his journey. Actually, Homer was not written by
Homer but by another man of that name.

Socrates was a famous Greek teacher who went around giving people
advice. They killed him. Socrates died from an overdose of wedlock.

In the Olympic Games, Greeks ran races, jumped, hurled the biscuits,
and threw the java. The reward to the victor was a coral wreath. The
government of Athen was democratic because the people took the law
into their own hands. There were no wars in Greece, as the mountains
were so high that they couldn't climb over to see what their neighbors
were doing. When they fought the Parisians, the Greeks were
outnumbered because the Persians had more men.

Eventually, the Ramons conquered the Geeks. History call people Romans
because they never stayed in one place for very long. At Roman
banquets, the guests wore garlic in their hair. Julius Caesar
extinguished himself on the battlefields of Gaul. The Ides of March
killed him because they thought he was going to be made king. Nero was
a cruel tyrany who would torture his poor subjects by playing the
fiddle to them.

Then came the Middle Ages. King Alfred conquered the Dames, King
Arthur lived in the Age of Shivery, King Harlod mustarded his troops
before the Battle of Hastings, Joan of Arc was cannonized by George
Bernard Shaw, and the victims of the Black Death grew boobs on their
necks. Finally, the Magna Carta provided that no free man should be
hanged twice for the same offense.

In midevil times most of the people were alliterate. The greatest
writer of the time was Chaucer, who wrote many poems and verse and
also wrote literature. Another tale tells of William Tell, who shot
an arrow through an apple while standing on his son's head.

The Renaissance was an age in which more individuals felt the value of
their human being. Martin Luther was nailed to the church door at
Wittenberg for selling papal indulgences. He died a horrible death,
being excommunicated by a bull. It was the painter Donatello's
interest in the female nude that made him the father of the
Renaissance. It was an age of great inventions and
discoveries. Gutenberg invented the Bible. Sir Walter Raleigh is a
historical figure because he invented cigarettes. Another important
invention was the circulation of blood. Sir Francis Drake circumcised
the world with a 100-foot clipper.

The government of England was a limited mockery. Henry VIII found
walking difficult because he had an abbess on his knee. Queen
Elizabeth was the ``Virgin Queen.'' As a queen she was a success. When
Elizabeth exposed herself before her troops, they all shouted
``hurrah.'' Then her navy went out and defeated the Spanish Armadillo.

The greatest writer of the Renaissance was William
Shakespear. Shakespear never made much money and is famous only
because of his plays. He lived in Windsor with his merry wives,
writing tragedies, comedies and errors. In one of Shakespear's famous
plays, Hamlet rations out his situation by relieving himself in a long
soliloquy. In another, Lady Macbeth tries to convince Macbeth to
kill the King by attacking his manhood. Romeo and Juliet are an
example of a heroic couplet. Writing at the same time as Shakespear
was Miquel Cervantes. He wrote ``Donkey Hote''. The next great author
was John Milton. Milton wrote ``Paradise Lost.'' Then his wife dies and
he wrote ``Paradise Regained.''

During the Renaissance America began. Christopher Columbus was a great
navigator who discovered America while cursing about the Atlantic. His
ships were called the Nina, the Pinta, and the Santa Fe. Later the
Pilgrims crossed the Ocean, and the was called the Pilgrim's
Progress. When they landed at Plymouth Rock, they were greeted by
Indians, who came down the hill rolling their was hoops before
them. The Indian squabs carried porposies on their back. Many of the
Indian heroes were killed, along with their cabooses, which proved
very fatal to them. The winter of 1620 was a hard one for the
settlers. Many people died and many babies were born. Captain John
Smith was responsible for all this.

One of the causes of the Revolutionary Wars was the English put tacks
in their tea. Also, the colonists would send their pacels through the
post without stamps. During the War, Red Coats and Paul Revere was
throwing balls over stone walls. The dogs were barking and the
peacocks crowing. Finally, the colonists won the War and no longer had
to pay for taxis.

Delegates from the original thirteen states formed the Contented
Congress. Thomas Jefferson, a Virgin, and Benjamin Franklin were two
singers of the Declaration of Independence. Franklin had gone to
Boston carrying all his clothes in his pocket and a loaf of bread
under each arm. He invented electricity by rubbing cats backwards
and declared ``a horse divided against itself cannot stand.'' Franklin
died in 1790 and is still dead.

George Washington married Matha Curtis and in due time became the
Father of Our Country. Them the Constitution of the United States was
adopted to secure domestic hostility. Under the Constitution the
people enjoyed the right to keep bare arms.

Abraham Lincoln became America's greatest Precedent. Lincoln's mother
died in infancy, and he was born in a log cabin which he built with
his own hands. When Lincoln was President, he wore only a tall silk
hat. He said, ``In onion there is strength.'' Abraham Lincoln write
the Gettysburg address while traveling from Washington to Gettysburg
on the back of an envelope. He also signed the Emasculation
Proclamation, and the Fourteenth Amendment gave the ex-Negroes
citizenship. But the Clue Clux Clan would torcher and lynch the
ex-Negroes and other innocent victims. On the night of April 14, 1865,
Lincoln went to the theater and got shot in his seat by one of the
actors in a moving picture show. The believed assinator was John
Wilkes Booth, a supposedly insane actor. This ruined Booth's career.

Meanwhile in Europe, the enlightenment was a reasonable time. Voltare
invented electricity and also wrote a book called ``Candy''. Gravity was
invented by Issac Walton. It is chiefly noticeable in the Autumn, when
the apples are flaling off the trees.

Bach was the most famous composer in the world, and so was
Handel. Handel was half German, half Italian and half English. He was
very large. Bach died from 1750 to the present. Beethoven wrote music
even though he was deaf. He was so deaf he wrote loud music. He took
long walks in the forest even when everyone was calling for
him. Beethoven expired in 1827 and later died for this.

France was in a very serious state. The French Revolution was
accomplished before it happened. The Marseillaise was the theme song
of the French Revolution, and it catapulted into Napoleon. During
the Napoleonic Wars, the crowned heads of Europe were trembling in
their shoes. Then the Spanish gorrilas came down from the hills and
nipped at Napoleon's flanks. Napoleon became ill with bladder problems
and was very tense and unrestrained. He wanted an heir to inheret his
power, but since Josephine was a baroness, she couldn't bear him any
children.

The sun never set on the British Empire because the British Empire is
in the East and the sun sets in the West. Queen Victoria was the
longest queen. She sat on a thorn for 63 years. He reclining years and
finally the end of her life were exemplatory of a great
personality. Her death was the final event which ended her reign.

The nineteenth century was a time of many great inventions and
thoughts. The invention of the steamboat caused a network of rivers to
spring up. Cyrus McCormick invented the McCormick Raper, which did the
work of a hundred men. Samuel Morse invented a code for
telepathy. Louis Pastuer discovered a cure for rabbis. Charles Darwin
was a naturailst who wrote the ``Organ of the Species''. Madman Curie
discovered radium. And Karl Marx became one of the Marx Brothers.

The First World War, cause by the assignation of the Arch-Duck by a
surf, ushered in a new error in the anals of human history.

\begin{flushright}
\textit{--- Richard Lederer, St.~Paul's School}
\end{flushright}